null_model

### 2020-07-03 analysis/ts_null_model.Rmd
refactor get_plot_FD function
save fd data for null model as csv, only include plots that meet the criteria for inclusion.

```{r}
library(tidyverse)
library(biodivTS)
```

@@

#drop trait columns that are the same value for every observation, and therefore providing no additional info
clean_trait_matrix <- function(trait_matrix){
  uniq_vals <- sapply(trait_matrix, function(x){length(unique(x))}) #get number of unique values of each trait
  col_names <- names(which(uniq_vals == 1)) #find columns where there is only one value for all observations
  return(select(trait_matrix, -col_names))
}

get_trait_matrix <- function(data, species_list, trait_list){ 
  traits <- data %>%
    filter(id %in% species_list) %>%
    dplyr::select(id, all_of(trait_list)) %>%
    distinct() %>%
    arrange(id) %>%
    column_to_rownames(var = "id")
}
```

@@

biotime_traits <- biotime_traits %>%
  mutate(plot_id = paste0(study_id, plot)) %>%
  filter(!is.na(id))

@@

sample_occurrence <- map_dfr(unique(biotime_traits$plot_id), get_plot_sample, data = biotime_traits, n = n) %>%
  replace(is.na(.), 0)

@@

get_plot_FD <- function(data, trait_range, occ_df, filter_plot) {
  sample_species_mat <- occ_df %>%
    filter(plot_id == filter_plot) %>%
    select(-c(richness, plot_id, n)) %>% #want only species x site info
    select_if(!colSums(.) %in% c(0, NA)) %>% #remove species that don't occur in the plot
    select(sort(tidyselect::peek_vars())) #sort columns to be in alphabetical order, required for dbFDI()

@@

  sample_fd <- as.data.frame(dbFD_joggle(x = sample_trait_mat, a = sample_species_mat, w.abun = FALSE))

@@

test <- get_plot_FD(biotime_traits, 17:64, sample_occurrence, 5617)
```

>>>>

```{r}
library(tidyverse)
library(biodivTS)
library(FD)
```

@@

#drop trait columns that are the same value for every observation, and therefore providing no additional info
clean_trait_matrix <- function(trait_matrix){
  # uniq_vals <- sapply(trait_matrix, function(x){length(unique(x))}) #get number of unique values of each trait
  # col_names <- names(which(uniq_vals == 1)) #find columns where there is only one value for all observations
  # return(select(trait_matrix, -col_names))
  trait_matrix %>%
  select_if(~ length(unique(.)) > 1) %>% #remove columns that have the same value for all columns
  mutate_at(vars(-id, -ends_with("5cat")), list(~scales::rescale(.,to=c(0, 1)))) %>% #rescale variables
  arrange(id) %>%
  column_to_rownames(var = "id")
}

get_trait_matrix <- function(data, species_list, trait_list){ 
  traits <- data %>%
    filter(id %in% species_list) %>%
    dplyr::select(id, all_of(trait_list)) %>%
    distinct()
}
```

@@

@@

#get the plots that meet our criteria for calculating FD
sample_plots <- biotime_traits %>%
  filter(!is.na(plot)) %>% #remove unlabeled plots 
  group_by(plot_id, year) %>%
  filter(n_distinct(id) > 4) %>%
  group_by(plot_id) %>%
  filter(n_distinct(year) > 4) %>% 
  ungroup()

sample_occurrence <- map_dfr(unique(sample_plots$plot_id), get_plot_sample, data = sample_plots, n = n) %>%
  replace(is.na(.), 0)

@@

get_plot_FD <- function(data, trait_range, occ_df, filter_plot) {
  plot_samples <- occ_df %>%
    filter(plot_id == filter_plot)

  sample_species_mat <- plot_samples %>%
    select(-c(richness, plot_id, n)) %>% #want only species x site info
    select_if(!colSums(.) %in% c(0, NA)) %>% #remove species that don't occur in the plot
    select(sort(tidyselect::peek_vars())) #sort columns to be in alphabetical order, required for dbFDI()

@@

  fs::dir_create(paste0(here::here("data"), "/fd_output"))

  #the convexhull c code creates an output file of vertices, in order to not have multiple threads writing
  #to the same file we have to create individual working directories
  tmp_dir <- paste0(here::here("data", "fd_output", "tmp"), "/tmp_", filter_plot)
  fs::dir_create(tmp_dir)
  setwd(tmp_dir)

  print(filter_plot)
  sample_fd <- as.data.frame(dbFD_joggle(x = sample_trait_mat, a = sample_species_mat, w.abun = FALSE)) %>%
    cbind(plot_samples %>% select(richness, plot_id, n))

  save_dir <- paste0(here::here("data", "fd_output", "null"))
  fs::dir_create(save_dir)
  write_tsv(sample_fd, paste0(save_dir, "/null", filter_plot, ".csv"))

@@

#test <- get_plot_FD(sample_plots, 18:65, sample_occurrence, 391)
null_FD <- map_dfr(unique(sample_plots$plot_id)[1:3], get_plot_FD, 
                   data = sample_plots, trait_range = 18:65, occ_df = sample_occurrence)
```
